---
title: "plot data"
author: ""
date: "April 2023"
output: 
  github_document:
    toc: true
    toc_depth: 2
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## setup

Here we load any required packages, source functions, and read in the processed meter readings data.

```{r set-up-read-meta, message=FALSE, warning=FALSE}

## load packages
library(dplyr)
library(lubridate)
library(ggplot2)

## source functions
source("functions_for_plots.R")  # currently empty!

## read meter data and check the structure
df <- read.csv("processed_meter_data.csv")
head(df)
```

## summarise and plot

Here we can start developing code to plot the data in various ways! 

Once we've finalised some plots, we can turn the code into functions with specific inputs and outputs, and put the code for those functions in `functions_for_plots.R`.


Here is a quick example of summarising the data.

```{r summaries-example}

## sum over all meters in a floor/location by year and month
summ1 <- df %>%
  dplyr::group_by(floor, location, year, month) %>%
  dplyr::summarise(total_usage = sum(both)) %>%   # both field = all lighting + power usage
  as.data.frame()

head(summ1)  # check output

```

Here is an example plot using the summary made above.

```{r plots-example}

## quick line graph of month usage data, split out by location field
ggplot2::ggplot(data = summ1, aes(x = month, y = total_usage, col = floor)) +
  ggplot2::geom_line() +
  ggplot2::facet_wrap(~location)


```

Continue from here!


